---
title: '"Vinho Verde" Wines Quality Modeling'
author: "The First Group (T.F.G.)"
date: "August 8, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

````{r}
Abstract: Wine classification is a difficult task since taste is the least understood of the human senses. In this research we propose a data mining approach to predict human wine taste preferences that is based on easily available analytical tests at the certification step. A large dataset is considered, with white and red vinho verde samples (from Portugal).  The support vector machine achieved promising results, outperforming the multiple regression and neural network methods. Such model is useful to support the oenologist wine tasting evaluations and improve wine production. Furthermore, similar techniques can help in target marketing by modeling consumer tastes from niche markets.

1 INTRODUCTION 
Data mining (DM) techniques aim at extracting knowledge from raw data. Several DM algorithms have been developed, each one with its own advantages and disadvantages (Witten and Frank, 2005). DM approaches have been applied to a large variety of problems, either for classification or regression. An interesting problem that has captured the attention of several researches is the prediction of wine quality (Cortez et al., 2009; Yin and Han, 2003). Wine industry is investing in new technologies for wine making and selling processes. A key issue in this context is wine certification which prevents the illegal adulteration and assures the wine quality. Wine certification is often assessed by physicochemical and sensory tests (Ebeler, 1999). The development of an accurate, computationally efficient and understandable prediction model can be of great utility for the wine industry. On the one hand, a good wine quality prediction can be very useful in the certification phase, since currently the sensory analysis is performed by human tasters, being clearly a subjective approach. An automatic predictive system can be integrated into a decision support system, helping the speed and quality of the oenologist performance. If it is concluded that several input variables are highly relevant to predict the wine quality, since in the production process some variables can be controlled, this information can be used to improve the wine quality. In this paper wine taste preferences are modelled by DM algorithms. 

```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

